\section{Classical Feature Extraction}
\subsection{Line Fitting}
\subsubsection*{Least Squares}
Fit $y=kx+m$ by minimizing $\sum_{i=1}^n(y_i-(kx_i+m))^2$.
Matrix form: $\min_{\boldsymbol{\theta}}\|A\boldsymbol{\theta}-\boldsymbol{y}\|_2^2$ with $A=[x_i,1]$, $\boldsymbol{\theta}=[k,m]^T$.
Solution: $\boldsymbol{\theta}=(A^TA)^{-1}A^T\boldsymbol{y}$. Unstable for near-vertical lines.

\subsubsection*{SVD Fitting}
Improved: $\min_{a,b,d}\sum(ax_i+by_i-d)^2$ s.t. $a^2+b^2+d^2=1$.
Matrix form: $\min_{\boldsymbol{h}}\|A\boldsymbol{h}\|^2$ with $A=[x_i,y_i,1]$, $\|\boldsymbol{h}\|=1$.
SVD: $A=UDV^T$. Solution: $\boldsymbol{h}=\pm\boldsymbol{v}_3$ (right singular vector for smallest singular value).

\subsubsection*{RANSAC}
Robust to outliers via random sampling. Steps: (1) choose $s$, $\delta$, $p$; (2) sample $s$ points; (3) fit model; (4) count inliers; (5) update best; (6) repeat $N$ times; (7) refit on best inliers.
Required iterations: $N=\frac{\log(1-p)}{\log(1-(1-e)^s)}$ where $e$ is outlier ratio.

\subsection{Corner Detection}
\subsubsection*{Harris Response}
Energy after shift $(u,v)$: $E(u,v)\approx[u,v]M[x_0,y_0]$
$[u,v]^T$.
Structure tensor: $M=\sum_{(x,y)\in N}$
$\begin{bmatrix}I_x^2&I_xI_y\\I_xI_y&I_y^2\end{bmatrix}$, often with Gaussian weighting.
Corner response: $\theta=\det M-\alpha(\operatorname{tr}M)^2=(I_x^2I_y^2-(I_xI_y)^2)-\alpha(I_x^2+I_y^2)^2$.
$\theta\gg0$: corner; $\theta\ll0$: edge; $\theta\approx0$: flat.

\subsubsection*{Harris Pipeline}
Steps: (1) compute gradients $I_x,I_y$; (2) compute $I_x^2,I_y^2,I_xI_y$; (3) Gaussian smooth the three maps; (4) compute $\theta$, threshold it; (5) NMS to keep local maxima.

\subsubsection*{Geometric Properties}
\textbf{Harris} has the following properties:
Translation equivariant: $T_{u,v}(\Theta(A))=\Theta(T_{u,v}(A))$.
Rotation equiv with \textbf{isotropic} filter: $R_\phi(\Theta(A))=\Theta(R_\phi(A))$.
Scale non-equiv: viewpoint changes break detection.

\subsection{Classical Recognition Pipeline}
\subsubsection*{Feature Detection}
Keypoint detector (Harris, DoG) + descriptor (SIFT). SIFT: 128-dim vector from $4\times4$ spatial grids with 8 orientation bins each, after orientation normalization.

\subsubsection*{Bag-of-Visual-Words}
Steps: (1) collect descriptors; (2) k-means cluster into $K$ visual words; (3) assign each descriptor to nearest word (hard/soft); (4) build $K$-dim frequency histogram.
Loses spatial information.

\subsubsection*{Limits of Classical CV}
Manual feature design; error accumulation across stages; weak high-level semantics; poor scaling with data; limited generalization.
