\section{Semantic Segmentation}
\subsection{Encoder-Decoder}
\subsubsection*{FCN Encoder}
Encoder $E_\theta:\mathcal{X}\to\mathcal{Z}$ downsamples by total factor $D=\prod s_\ell$. Bottleneck size $H_L=\lfloor H_0/D\rfloor$, $W_L=\lfloor W_0/D\rfloor$.
Decoder $D_\phi$ upsamples back to original size. Segmentation output: $\hat y\in\mathbb{R}^{K\times H_0\times W_0}$.
Seg loss: pixel-wise cross-entropy over $K$ classes.

\subsubsection*{Non-learnable Upsampling}
Bilinear: interpolate from source coordinates.
Nearest neighbor: replicate each pixel to $r\times r$ block.
Max unpooling: record max positions during pooling; place values back, fill others with 0.

\subsubsection*{Transposed Convolution}
Equivalent to conv after zero-insertion and padding. Output size: $h_{out}'=s(h_{in}'-1)+a+k-2p$.
Checkerboard artifact when $k$ not divisible by $s$.
Alternatives: resize-conv (upsample then $3\times3$ conv), PixelShuffle.

\subsection{UNet}
\subsubsection*{Skip Connections}
Concatenate encoder features to decoder at same resolution. Bottleneck handles global context; skip links provide per-pixel spatial detail.

\subsubsection*{Information Split}
Encoder-Decoder alone forces bottleneck to carry both context and spatial info. Skip links split the burden, improving optimization.

\subsection{Segmentation Evaluation}
\subsubsection*{Overlap Metrics}
IoU$=\frac{TP}{TP+FP+FN}=\frac{|P\cap G|}{|P\cup G|}$.
mIoU$=\frac{1}{K}\sum_{k=1}^K\text{IoU}_k$ (fair for class imbalance).
Pixel accuracy can be misleading when positives are rare.

\subsubsection*{Soft IoU Loss}
Soft Intersection$=\sum_i p_i y_i$; Soft Union$=\sum_i(p_i+y_i-p_iy_i)$.
$\mathcal{L}_{\text{IoU}}=1-\frac{\sum_i p_iy_i+\epsilon}{\sum_i p_i+\sum_i y_i-\sum_i p_iy_i+\epsilon}$.
Multi-class: average per-class soft IoU loss.
