\section{Object Detection}
\subsection{Basics}
\subsubsection*{Task Definition}
Output axis-aligned bbox $(x,y,h,w)$ plus class confidence score. Backbone CNN extracts feature map.

\subsubsection*{Regression Losses}
L1: $\ell=\sum|\Delta_i|$ (robust, constant gradient). L2: $\ell=\sum\Delta_i^2$ (sensitive to outliers).
Smooth L1 (Huber): $\text{smooth}_{L1}(x)=0.5x^2$ if $|x|<1$, else $|x|-0.5$. Best of both.

\subsection{Two-Stage Detectors}
\subsubsection*{R-CNN to Fast R-CNN}
R-CNN: selective search proposals, each cropped and fed to CNN independently. Slow, not end-to-end, loses context.
Fast R-CNN: CNN first on whole image, then crop feature map (RoI Pool). Faster but still uses selective search.

\subsubsection*{Faster R-CNN}
RPN replaces selective search on shared feature map. Pipeline: Backbone $\to$ RPN proposals $\to$ RoI Pool/Align $\to$ cls + reg heads.

\subsubsection*{RoI Generation}
RPN outputs proposals with objectness scores. Post-process: NMS ($\tau=0.7$), top-K, positive/negative sampling by IoU thresholds (pos $>0.7$, neg $<0.3$, ignore middle).
RPN anchor regression targets: $t_x=\frac{x_g-x_a}{w_a}$, $t_y=\frac{y_g-y_a}{h_a}$, $t_w=\log\frac{w_g}{w_a}$, $t_h=\log\frac{h_g}{h_a}$.
RPN loss: $L=L_{\text{cls}}+\lambda L_{\text{reg}}$ with BCE for cls and Smooth L1 for reg (positive only).

\subsection{Single-Stage Detectors}
\subsubsection*{Core Mechanism}
YOLO/SSD: dense prediction on feature map grid $H\times W$ with $B$ anchors per cell. Output $H\times W\times(5B+K)$. Direct cls + reg without proposals.

\subsubsection*{Training Design}
Category-specific (vs RPN's foreground/background). Faster inference, slightly lower accuracy historically.

\subsection{Post-processing and Evaluation}
\subsubsection*{Non-Maximum Suppression}
NMS: sort by score; greedily keep highest, remove IoU$>\tau$. Soft-NMS: decay scores by overlap instead of hard removal. Class-wise NMS: run per category.

\subsubsection*{Detection Metrics}
TP: correct class + IoU$\ge\tau$ + unmatched GT. FP: wrong class or IoU$<\tau$ or duplicate match. FN: unmatched GT.
AP: area under PR curve (11-point interpolation). mAP: mean over classes and/or IoU thresholds.
